# Project : Heart Disease Prediction
# Required Libraries
```{r}
library(dplyr)
library(corrplot)
library(ggplot2)
library(interactions)
library(caTools)
library(MASS)
library(car)
library(aod)
```

## Data Overview

```{r}
heart <- read.csv('heart.csv')
heart$age <- heart$ï..age

heart <- heart[-c(1)]

cat_vars <- c( 'cp', 'slope', 'ca', 'thal','sex','exang','fbs', 'target', 'restecg')
num_vars <- colnames(heart)[!colnames(heart) %in% cat_vars]


heart[cat_vars] <- lapply(heart[cat_vars], as.factor)
summary(heart)
```
- in ca : The value of 4 is invalid as data description says that the value for ca lies between  0 to 3.

- thal : 0 is an invalid value as it can either be 1 that is Normal, 1 Fixed defect, 2 Reversible Defect.

Angina is the main symptom of CHD. Angina is a chest pain. SO if chest pain is Angina it is a sure case of CHD. In case of asymptomatic and non-anginal pain, analysis of other variables is needed.



# Dropping invalid values from ca and thal



```{r}
heart <- subset(heart, ca!=4)
heart <- subset(heart, thal!=0)
heart$ca<- droplevels(heart$ca)
heart$thal<- droplevels(heart$thal)

```



## Analyzing Dataset:

```{r}
sum(is.na(heart))
```
No Missing Data

Checking Duplicates

```{r}
dim(distinct(heart))

heart <- unique(heart)
```
1 Duplicate row found, Number of observations are now 302.



# Dropping invalid values from ca and thal
```{r}
heart <- subset(heart, ca!=4)
heart <- subset(heart, thal!=0)
heart$ca<- droplevels(heart$ca)
heart$thal<- droplevels(heart$thal)
```
```{r}
str(heart)
```


## Numerical Variables : Two Sample T-Test

```{r}
# Variance  Check
var.test(subset(heart, target==0)$trestbps, subset(heart, target==1)$trestbps)
var.test(subset(heart, target==0)$chol, subset(heart, target==1)$chol)
var.test(subset(heart, target==0)$thalach, subset(heart, target==1)$thalach)
var.test(subset(heart, target==0)$oldpeak, subset(heart, target==1)$oldpeak)
var.test(subset(heart, target==0)$age, subset(heart, target==1)$age)
```
variance equal for trestbps and chol, for rest variances are unequal

```{r}
t.test(subset(heart, target==0)$trestbps, subset(heart, target==1)$trestbps, var.equal = TRUE)
t.test(subset(heart, target==0)$chol, subset(heart, target==1)$chol, var.equal = TRUE)
t.test(subset(heart, target==0)$thalach, subset(heart, target==1)$thalach, var.equal = FALSE)
t.test(subset(heart, target==0)$oldpeak, subset(heart, target==1)$oldpeak, var.equal = FALSE)
t.test(subset(heart, target==0)$age, subset(heart, target==1)$age, var.equal = FALSE)
```
All are significant variables except chol.

## Chi-Sq Test

```{r}
table(heart$sex, heart$target)
chisq.test(x=heart$sex, y=heart$target)
```
significant

```{r}
table(heart$cp, heart$target)
chisq.test(x=heart$cp, y=heart$target)

table(heart$fbs, heart$target)
chisq.test(x=heart$fbs, y=heart$target)

table(heart$exang, heart$target)
chisq.test(x=heart$exang, y=heart$target)


```
significant 

```{r}
# table(heart$fbs, heart$target)
# chisq.test(x=heart$fbs, y=heart$target)
```
not significant  -fbs
  
```{r}
table(heart$restecg, heart$target)
chisq.test(x=heart$restecg, y=heart$target)
```
Since Type-2 has just 4 observations, we will megre it with type-1 and change the levels to 0 being normal and 1 being abnormal
```{r}
heart$restecg[heart$restecg == 2] <- 1
heart$restecg <- droplevels(heart$restecg)
table(heart$restecg, heart$target)
chisq.test(x=heart$restecg, y=heart$target)
```
significant

```{r}
# table(heart$exang, heart$target)
# chisq.test(x=heart$exang, y=heart$target)
```
```{r}
# table(heart$slope, heart$target)
# chisq.test(x=heart$slope, y=heart$target)
```
significant

```{r}
table(heart$ca, heart$target)
chisq.test(x=heart$ca, y=heart$target)
```

significant

```{r}
table(heart$thal, heart$target)
chisq.test(x=heart$thal, y=heart$target)
```


significant


#chose initial variables -- remove col and fbs due to insignificance in either T test or chi square test

```{r}
heart_v1 <- subset(heart, select = -c(chol,fbs))
num_vars_v1 <- num_vars[-2]
cat_vars_v1 <- cat_vars[-7]
heart_target_n <- ifelse(heart$target == 0, 0, 1)
```




## Check for Multicolinearity between predictors


### Numerical Variables

```{r}
coeff <- cor(heart_v1[num_vars_v1])
coeff

```

```{r}
corrplot(coeff, type="upper", order="hclust", tl.col="red", tl.srt=45, diag = FALSE, method = 'number',
         col= colorRampPalette(c("blue","green", "yellow"))(10))
```

```{r}
my_cols <- c("#00AFBB", "#E7B800")  #, "#FC4E07"
pairs(heart_v1[num_vars_v1], pch = 19,
      col = my_cols[heart_v1$target],
      lower.panel=NULL)

```  

No high correlation in variables, so no multicollinearity between them



## Interaction Terms 
Just pick one interaction if there is more than 1
Even if there is no interaction, pick one to compare with and without interaction model



```{r}

fit1 <- glm(data = heart_v1, formula = target ~ sex * ca, family = 'binomial')
cat_plot(fit1,pred = ca, modx = sex, geom = 'line')

fit2 <- glm(data = heart_v1, formula = target ~ sex * thal, family = 'binomial')
cat_plot(fit2,pred = thal, modx = sex, geom = 'line')

fit3 <- glm(data = heart_v1, formula = target ~ sex * slope, family = 'binomial')
cat_plot(fit3,pred = slope, modx = sex, geom = 'line')

interaction.plot(heart_v1$ca,heart_v1$sex,as.numeric(heart_v1$target))


```

## Split the dataset : Train & Testset

```{r}
set.seed(0)
data_split <- sample.split(heart_v1$target, SplitRatio = 0.7)
heart_v1.train <- subset(heart_v1, data_split == T)
heart_v1.test <- subset(heart_v1, data_split == F)
heart_target_n.test <- subset(heart_target_n, data_split == F)
nrow(heart_v1.test)
nrow(heart_v1.train)

```

##Stepwise Variable selection

### Without Interaction
```{r}
full_fit_wo_int <- glm(data = heart_v1.train, family = binomial, formula = target~.)
stepAIC(full_fit_wo_int, direction = "both", trace =TRUE)
```

### With Interaction : Sex * Thal
```{r}
full_fit_int <- glm(data = heart_v1.train, family = binomial, formula = target~cp+trestbps+restecg+thalach+exang+oldpeak+slope+ca+age+sex+thal+sex:thal)
stepAIC(full_fit_int, direction = "both", trace =TRUE)
```

Without Interaction Model : target ~ sex + cp + slope + ca + thal
Wih Interaction model : target ~ sex + cp + slope + ca + thal + sex:thal

## Compare Models using 
### Log-likelihood Ratio Test : Use Anova to check goodness of fit
```{r}
fit_wo_int <- glm(formula = target ~ cp + slope + ca + sex + thal, family = binomial, 
    data = heart_v1.train)
fit_int <- glm(formula = target ~ cp + slope + ca + sex + thal+ sex:thal, family = binomial, 
    data = heart_v1.train)
summary(fit_wo_int)
summary(fit_int)
```
### Checking significance of parameter in model : Interaction Term
LRT : 
```{r}
res_wo_int <- 112.19
res_int <- 111.24
diff <- res_wo_int - res_int
pchisq(diff, df = 2, lower.tail = FALSE)
anova(fit_wo_int, fit_int, test = "LRT")
```
Since, p-value is high, so we consider model without interaction.

### Wald Test
From results, p-value of interaction term for z-test statistic is greater than 0.05 indicating acceptance of null hypothesis, i.e., beta = 0 or no intercation term

```{r}
wald.test(b = coef(fit_int), Sigma = vcov(fit_int), Terms = 13:14)
```


## Final Data Model Training

## Predicting Responses
```{r}
heart_v1.test$predicted <- predict(fit_wo_int, newdata = heart_v1.test, type = "response")
```


# Classification Report(Sensitivity, Specificity, Accuracy)

## Evaluating cut-off probability
```{r}
pi0 <- sum(heart_target_n)/nrow(heart_v1)
pi0
```
Taking cut-off probability as 0.54 

## Comapre a pre-defined pi value with probablities obtained.
```{r}
heart_v1.test$predicted <- ifelse(heart_v1.test$predicted > pi0, 1, 0)
heart_v1.test$predicted
```


## Create Confusion matrix for accuracy
```{r}
library(caret)
Actual <- heart_target_n.test
Predicted <- heart_v1.test$predicted
confusionMatrix(as.factor(Predicted),as.factor(Actual),positive = "1")
```

## Acuracy for Model with interaction

```{r}
heart_v1.test$predicted_int <- predict(fit_int, newdata = heart_v1.test, type = "response")
heart_v1.test$predicted_int <- ifelse(heart_v1.test$predicted_int > pi0, 1, 0)
Predicted_int <- heart_v1.test$predicted_int
confusionMatrix(as.factor(Predicted_int),as.factor(Actual), positive="1")
```

## ROC curve 
```{r}
library(PRROC)

PRROC_obj <- roc.curve(scores.class0 = Predicted, weights.class0=Actual,
                       curve=TRUE)
plot(PRROC_obj)
```

```{r}
PRROC_obj_int <- roc.curve(scores.class0 = Predicted_int, weights.class0=Actual,
                       curve=TRUE)
plot(PRROC_obj_int)
```

## Lack of Fit Test (Homser-Lemshow Test-ungrouped data)

```{r}
library(ResourceSelection)
hoslem.test(fit_wo_int$y, fitted(fit_wo_int), g=10)
```

```{r}
hoslem.test(fit_int$y, fitted(fit_int), g=10)
```





